### 2016-01-11 papers/CAFe/HIPS-paper/intro.tex
Added a couple of paragraphs on OCR.

With the addition of coarrays in the 2008 standard, Fortran is now a \emph{parallel}
language.  Originally called Coarray Fortran (CAF) by its
authors\cite{Numrich:1998:CFP:289918.289920}, parallism in Fortran is similar to MPI in
that all cooperating CAF processes (called \emph{images}) execute a Single Program but with
different (Multiple) Data (SPMD).  Key concepts in CAF are parallel execution, distributed
memory allocation, and remote memory transfer.  Parallel execution and memory allocation
are straightforward as each program is replicated $NP$ times with each program image
allocating its own memory locally.  Special syntax is provided to indicate which arrays
are visible between distributed memory processes and remote memory transfer is indicated
with new array syntax using square brackets, for example,
\small
\begin{verbatim}
   U(1) = U(1)[3]
\end{verbatim}
\normalsize
copies memory from the first element of \texttt{U} residing on image number 3 into
the corresponding element of the \emph{local} array \texttt{U}.  The fundamental programming
model is similar to that provided by message passing systems.

@@

Historically it has been hard to create parallel programs.  The
responsibility (and difficulty) of creating \emph{correct} parallel
programs can be viewed as being spread between the programmer and the
compiler.  Ideally we would like to have parallel languages that make
it easy for the programmer to express correct parallel programs --- and
conversely it should be difficult to express incorrect parallel
programs.  Unfortunately, many current languages and standards place
all of the responsibility on the user.  The best example of this are
programs written using MPI (Message Passing Interface), where the
programmer expresses all parallelism in terms of calls to library
routines and the serial C or Fortran compiler knows nothing about the
parallel semantics of the combined language plus MPI library.

@@

>>>>

Another option is to abandon the stricture of message passing within a procedural
programming language altogether.  The Open Community Runtime (OCR) project provides an
asynchronous task-based runtime designed for machines with high-core count, such as can be
expected in any exascale hardware design\cite{OCR:SC:2012}.  The project's goal is to
create a framework and reference implementation to help developers explore programming
methods to improve the power efficiency, programmability, and reliability of HPC
applications while maintaining application performance.

OCR helps the application developer with the complex process of writing multi-core
applications.  It does this by providing facilities for creating and executing a set of
tasks.  The OCR runtime schedules the execution of these tasks based on events which
embody dataflow and code flow dependencies; when all of the dependencies for a task are
satisfied, the task can be run.

However, it cannot be emphasized too strongly that like using the OpenCL programming
environment, actually creating real scientific applications based on OCR is a high hurdle
for the application developer.  A goal of this paper is to examine ways in
which high-level syntax based on Fortran coarrays can be utilized by compiler-based tools
to \emph{target} programming environments like OpenCL or OCR --- thus freeing the developer from
the task of frequently porting the application to new complex environments.  The hope is
that high-level syntax can hide the complexities of the underlying software and hardware.

With the addition of coarrays in the 2008 standard, Fortran is now a \emph{parallel}
language.  Originally called Coarray Fortran (CAF) by its
authors\cite{Numrich:1998:CFP:289918.289920}, parallism in Fortran is similar to MPI in
that all cooperating CAF processes (called \emph{images}) execute a Single Program but with
different (Multiple) Data (SPMD).  Key concepts in CAF are parallel execution, distributed
memory allocation, and remote memory transfer.  Parallel execution and memory allocation
are straightforward as each program is replicated $NP$ times with each program image
allocating its own memory locally.  Special syntax is provided to indicate which arrays
are visible between distributed memory processes and remote memory transfer is indicated
with new array syntax using square brackets, for example,
%\small
\begin{verbatim}
   U(1) = U(1)[3]
\end{verbatim}
%\normalsize
copies memory from the first element of \texttt{U} residing on image number 3 into
the corresponding element of the \emph{local} array \texttt{U}.  The fundamental programming
model is similar to that provided by message passing systems.

@@

\begin{comment}
Historically it has been hard to create parallel programs.  The
responsibility (and difficulty) of creating \emph{correct} parallel
programs can be viewed as being spread between the programmer and the
compiler.  Ideally we would like to have parallel languages that make
it easy for the programmer to express correct parallel programs --- and
conversely it should be difficult to express incorrect parallel
programs.  Unfortunately, many current languages and standards place
all of the responsibility on the user.  The best example of this are
programs written using MPI (Message Passing Interface), where the
programmer expresses all parallelism in terms of calls to library
routines and the serial C or Fortran compiler knows nothing about the
parallel semantics of the combined language plus MPI library.

@@

\end{comment}
